---
title: "In-class Exercise 1: My first date with Geospatial Data Analytics."
date: "18 Nov 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Getting started

The code chunk below load the following packages:

-   tmap: for thematic mapping

-   sf: for geospatial data handling

-   tidyverse: for non-spatial data handling

```{r}
pacman::p_load(tmap,sf,tidyverse)
```

## Preparing the Flow data

### Importing the OD data

Firstly, we will import the *Passenger Volume by Origin Destination* data set downloaded from LTA DataMall by using `read_csv()` of **readr** package.

```{r}
odbus <-read_csv("data/aspatial/origin_destination_bus_202308.csv")
```

A quick check of odbus tibble data frame

```{r}
odbus$ORIGIN_PT_CODE<-as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE<-as.factor(odbus$DESTINATION_PT_CODE)
```

### Extracting the study data

For the purpose of this exercise, we will extract commuting flows on weekday and between 7 and 9 o'clock

```{r}
origtrip_7_9<-odbus %>% 
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
busstop <- st_read(dsn = "data/geospatial", layer = "BusStop")%>%
  st_transform(crs=3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MPSZ-2019") %>%
  st_transform(crs=3414)
mpsz
```
