---
title: "Hands-on Exercise 1"
editor: visual
---

## Overview

This is the overview paragraph

## Getting Started

This is the getting start paragraph

## Importing Geospatial data

Importing MP14 subzone

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

Importing cycling path

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")
```

Import preschools

```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

## Checking content

```{r}
st_geometry(mpsz)
```

Taking a glimpse at the data

```{r}
glimpse(mpsz)
```

Using *head()* to take a look at first *n* rows

```{r}
head(mpsz, n=5)
```

## Plotting the Geospatial Data

```{r}
plot(mpsz)
```

Showing default plot

```{r}
plot(st_geometry(mpsz))
```

Showing by specific attribute

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Map Projection Transformation

```{r}
st_crs(mpsz)
```

Assigning correct EPSG code

```{r}
mpsz3414<-st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
```

Transforming projection of preschool from wgs84 to svy21. This is to have all of the coordinates be using the same coordinate system.

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
preschool3414
```

## Importing aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
list(listings)
```

Creating dataframe from aspatial dataframe

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
glimpse(listings_sf)
```

## Geoprocessing

Determining the amount of land needed and the total area

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

Find number of pre-schools in each planning subzone

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
mpsz3414
```
## Explorotary Data Analysis(EDA)
Plot preschool density with histogram
```{r}
hist(mpsz3414$`PreSch Density`)
```
```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```
Scatter plot Preschool density vs Preschool count
```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```